Galaxy Zoo: Morphological Classifications for 150,000 Galaxies in HST Legacy Imaging

### 2016-05-07 writeup/gz_hubble_data.tex
Text on interface and classification stats

@@

{\note Describe the GZ interface and the hierarchical decision tree.}

\begin{figure*}
\center
\includegraphics[width=160mm]{figures/gzh_interface.png}
\caption{Screenshot of the Galaxy Zoo: Hubble interface at the beginning of a classification, with the user ready to select an answer for the first question in the decision tree.\label{fig:interface}}
\end{figure*}

@@

In the \ferengi{} data, these numbers severely limit the amount of information that can be extracted for the higher-tier questions. The debiasing technique used (Section~\ref{ssec:zeta}) requires that at least 10~users answer each question for a galaxy with $z_\mathrm{sim}=0.3$ \emph{and} the corresponding image at higher redshift. This requirement is (by default) met by all galaxies for the smooth/features question. However, this is often \emph{not} met for questions beyond Task~01; on average. $60\%\pm24\%$ of the galaxies do not have sufficient data to measure a correction, as compared to 2.0\% achieved for Task~01 (Table~\ref{tbl:ferengi_corrections}). Of the remaining \ferengi{} galaxies which did meet this requirement, polynomials were fit to the data in each redshift-surface brightness bin, as was done for Task~01. The goodness-of-fit was evaluated in each bin, for each Task, using a normalized $\chi^2$ metric. The mean $\chi^2$ for all higher-order tasks was $0.10\pm0.04$, significantly larger than the Task 01 value of $0.04$. Based on this statistic, the data cannot not be modeled with sufficient confidence for the higher order tasks. An example of the debiasing technique applied to $p_{\rm bar}$ is shown in Section~\ref{sec:ferengi_bar}; visual inspection of Figure~\ref{fig:p_vs_p_bar} confirms that the points in each bin are scattered to a degree at which no strong correlation is present. For these reasons, debiased vote fractions are only offered in the GZH catalog for Task~01 (smooth/features).

@@

In GZH, volunteers answer questions about a galaxy's morphology in a decision tree format. With this structure, questions shown to a user are dependent on their answers to the previous questions. For example, a user is only asked Task 11, {\it ``How many spiral arms?''} if they had answered ``yes'' to Task 10, which asks whether any spiral arms are present. This structure is used to reduce the workload of the users by only asking questions relevant to the particular galaxy; in this example, it would not be useful, nor would it make physical sense, to ask a user how many spiral arms in a galaxy which they had declared had no spiral arms to begin with. Figure~\ref{fig:decisiontree} shows the possible paths available for answering the questions offered in GZH. The colors represent the tier level of Tasks, which indicate the number of previous Tasks the given Task is dependent upon. The arm-number question, Task 11, is a 5th tier Task, meaning that whether this question is seen by a user is dependent on four Tasks preceding it. In this case, it is only shown to volunteers who voted the galaxy was featured/disk-shaped in Task 01, not clumpy in Task 12, not edge-on in Task 02, and had spiral structure in Task 10.

@@

Tasks 03, 04 and 05 have an additional possible pathway; as shown in Figure~\ref{fig:decisiontree}, a user might also be shown this question if they select ``featured/disk'' in Task 01, ``clumpy'' in Task 12, two or more clumps in Task 16, and ``spiral arrangement'' in Task 15. After applying the appropriate thresholds for this path, $< 0.5\%$ of the galaxies which have $\ge 20$ answers to these questions used this pathway to arrive at these Tasks. Further, of these subjects, none were found to exhibit disk structure, although the clumps within were arranged in a spiral pattern.

@@

For galaxies identified as ``clumpy'' in GZH, the number of clumps can be determined using Task 16. Table~\ref{tbl:thresholds} can be used to select 8,444 galaxies measured as ``clumpy'' using $p_{\rm features} > 0.23$ and $p_{\rm clumpy,yes}>0.80$ to ensure the vote fractions for Task 16 are well-sampled. The clump number can be reasonably identified for 1,112 of the clumpy galaxies; for the remainder, the unique clumps were less distinguishable from each other and the exact number of clumps could not be deduced without careful visual inspection. In the 1,112 which did have distinguishable clumps, there are 61 one-clump galaxies using $p_{\rm 1~clump}>0.50$, 442 two-clump galaxies using $p_{\rm 2~clumps}>0.80$, 275 three-clump galaxies using $p_{\rm 3~clumps}>0.75$, 71 four-clump galaxies using $p_{\rm 4~clumps} > 0.70$, and 263 galaxies with more than four clumps using $p_{\rm >4~clumps}>0.70$. Alternatively, these data may be used to create more general samples of clumpy galaxies with few clumps and many clumps. A sample of 989 ``few clumps'' galaxies can be made using $(p_{\rm 1~clump} + p_{\rm 2~clumps}) > 0.5$ and 2,910 ``many clumps'' galaxies using $(p_{\rm 3~clumps}+p_{\rm 4~clumps}+p_{\rm >4~clumps})>0.5$.

>>>>

Classifications for a particular user are aggregated and weighted only if they were logged into GZH under their username (which is encouraged by not required for participation). Classifications by users not logged in are marked as ``Anonymous''; no weighting is applied to these classifications, but they are otherwise treated identically in the data processing.

\begin{figure*}
\center
\includegraphics[width=160mm]{figures/gzh_interface.png}
\caption{Screenshot of the Galaxy Zoo: Hubble interface (\url{http:/zoo3.galaxyzoo.org}) at the beginning of a classification, with the user ready to select an answer for the first question in the decision tree.\label{fig:interface}}
\end{figure*}

Classifications for GZH are made using a web-based interface (Figure~\ref{fig:interface}), similar in design to Galaxy~Zoo and Galaxy~Zoo~2. The front-end runs on a Ruby~on~Rails framework with classifications stored in a MySQL backend. Users are shown a randomly-selected color composite image from the GZH sample; the default shows the image with a black sky background, although users have the options to invert the color palette if desired. The questions and responses for morphology appear on the right side of the image as a panel, including both text and icons. There is no tutorial required for participation, although users can access an extensive ``Help'' section containing example images and descriptive text for all the morphological labels.

@@

The procedure for classifying an image in GZH follows a hierarchical decision tree (Figure~\ref{fig:decisiontree}). Every classification begins with the user identifying whether the object at the center of the image is a smooth galaxy, a galaxy with a disk or other features, or a star/artifact. Subsequent questions in the tree depend on the previous answer(s) given by the user; the structure of the decision tree is designed so that the user will answer every question relevant to the morphology that they have identified. Questions that are not answered are implicitly assumed to be absent in the image --- for example, if the user identifies a galaxy as being smooth, they are not asked to count the number of spiral arms. For every task, the user chooses a single answer before continuing to the next question; they have also the option to restart a classification in-progress if a mistake has been made.

The GZH decision tree has four broad sets of morphologies classified by the users. The first option is designating the object as a star or artifact (the result of either bad data or incorrect classification by the ACS pipeline as a galaxy); in this case, the classification ends and no further questions are asked. The second set are smooth galaxies, intended to identify ellipticals/early-types; users also measure the relative axial ratio (roundness) for these galaxies. The third set is for disk/late-type galaxies, which measures the features necessary to place a galaxy on the standard Hubble tuning fork (bars, spiral arm, strength of the central bulge). The final set, which is new to the high-redshift targets in GZH, asks users to identify objects dominated by clumpy morphologies. Further annotation for these galaxies includes assessing the number, arrangement, relative brightness, and location of the clumps within the galaxy. Finally, every classification of a galaxy gives users the option of annotating ``odd'' features within the image; these labels are typically rare in large surveys, including dust lanes, gravitational lenses, and major mergers.

The number of independent classifications per subject collected by GZH is on average higher than GZ1 or GZ2, due to both the increased complexity of the decision tree and the relative difficulty of classifying images of small and distant galaxies. Images from the main AEGIS, GEMS, and GOODS data sets have a median of 122~independent classifications per image. The remainder of the images were set slightly lower, either from having a later activation date (COSMOS, simulated AGN) or by setting a lower retirement rate (the low-redshift SDSS Stripe~82 galaxies). Images from these samples have a median of 46-48~classifications per image (Figure~\ref{fig:classification_hist}).

\begin{figure}
\center
\includegraphics[width=0.5\textwidth]{figures/classification_hist.pdf}
\caption{Distribution of the total number of classifications per image for GZH, split by the survey types.}
\label{fig:classification_hist}
\end{figure}

The GZH project was launched on 23~Apr~2010 with the inclusion of the AEGIS, GEMS, GOODS 2-epoch, and SDSS Stripe~82 images. Images from COSMOS and the simulated AGN were activated in Dec 2010, as well as a small sample of AEGIS/GEMS/GOODS images mistakenly excluded from the original sample based on an overly restrictive cut on blended and/or saturated objects. The main site operated until the launch of the Galaxy~Zoo:~CANDELS project on 10~Sep~2012. Classifications for the GOODS 5-epoch images were obtained with the fourth version of the site\footnote{\url{http://www.galaxyzoo.org}} from Mar-Jun 2015. The GZH project had a total of 10,349,357~classifications from 93,898 registered participants. % Sum is the total of the GZH SQL table plus the GOODS full-depth imaging in GZ4.

@@

In the \ferengi{} data, these numbers severely limit the amount of information that can be extracted for the higher-tier questions. The debiasing technique used (Section~\ref{ssec:zeta}) requires that at least 10~users answer each question for a galaxy with $z_\mathrm{sim}=0.3$ \emph{and} the corresponding image at higher redshift. This requirement is (by default) met by all galaxies for the smooth/features question. However, this is often \emph{not} met for questions beyond Task~01; on average. $60\%\pm24\%$ of the galaxies do not have sufficient data to measure a correction, as compared to 2.0\% achieved for Task~01 (Table~\ref{tbl:ferengi_corrections}). Of the remaining \ferengi{} galaxies which did meet this requirement, polynomials were fit to the data in each redshift-surface brightness bin, as was done for Task~01. The goodness-of-fit was evaluated in each bin, for each Task, using a normalized $\chi^2$ metric. The mean $\chi^2$ for all higher-order tasks was $0.10\pm0.04$, significantly larger than the Task~01 value of $0.04$. Based on this statistic, the data cannot not be modeled with sufficient confidence for the higher order tasks. An example of the debiasing technique applied to $p_{\rm bar}$ is shown in Section~\ref{sec:ferengi_bar}; visual inspection of Figure~\ref{fig:p_vs_p_bar} confirms that the points in each bin are scattered to a degree at which no strong correlation is present. For these reasons, debiased vote fractions are only offered in the GZH catalog for Task~01 (smooth/features).

@@

In GZH, volunteers answer questions about a galaxy's morphology in a decision tree format. With this structure, questions shown to a user are dependent on their answers to the previous questions. For example, a user is only asked Task~11, {\it ``How many spiral arms?''} if they had answered ``yes'' to Task~10, which asks whether any spiral arms are present. This structure is used to reduce the workload of the users by only asking questions relevant to the particular galaxy; in this example, it would not be useful, nor would it make physical sense, to ask a user how many spiral arms in a galaxy which they had declared had no spiral arms to begin with. Figure~\ref{fig:decisiontree} shows the possible paths available for answering the questions offered in GZH. The colors represent the tier level of Tasks, which indicate the number of previous Tasks the given Task is dependent upon. The arm-number question, Task~11, is a 5$^\mathrm{th}$-tier Task, meaning that whether this question is seen by a user is dependent on four Tasks preceding it. In this case, it is only shown to volunteers who voted the galaxy was featured/disk-shaped in Task~01, not clumpy in Task~12, not edge-on in Task~02, and had spiral structure in Task~10.

@@

Tasks 03, 04 and 05 have an additional possible pathway; as shown in Figure~\ref{fig:decisiontree}, a user might also be shown this question if they select ``featured/disk'' in Task~01, ``clumpy'' in Task~12, two or more clumps in Task~16, and ``spiral arrangement'' in Task~15. After applying the appropriate thresholds for this path, $< 0.5\%$ of the galaxies which have $\ge 20$ answers to these questions used this pathway to arrive at these Tasks. Further, of these subjects, none were found to exhibit disk structure, although the clumps within were arranged in a spiral pattern.

@@

For galaxies identified as ``clumpy'' in GZH, the number of clumps can be determined using Task~16. Table~\ref{tbl:thresholds} can be used to select 8,444 galaxies measured as ``clumpy'' using $p_{\rm features} > 0.23$ and $p_{\rm clumpy,yes}>0.80$ to ensure the vote fractions for Task~16 are well-sampled. The clump number can be reasonably identified for 1,112 of the clumpy galaxies; for the remainder, the unique clumps were less distinguishable from each other and the exact number of clumps could not be deduced without careful visual inspection. In the 1,112 which did have distinguishable clumps, there are 61 one-clump galaxies using $p_{\rm 1~clump}>0.50$, 442 two-clump galaxies using $p_{\rm 2~clumps}>0.80$, 275 three-clump galaxies using $p_{\rm 3~clumps}>0.75$, 71 four-clump galaxies using $p_{\rm 4~clumps} > 0.70$, and 263 galaxies with more than four clumps using $p_{\rm >4~clumps}>0.70$. Alternatively, these data may be used to create more general samples of clumpy galaxies with few clumps and many clumps. A sample of 989 ``few clumps'' galaxies can be made using $(p_{\rm 1~clump} + p_{\rm 2~clumps}) > 0.5$ and 2,910 ``many clumps'' galaxies using $(p_{\rm 3~clumps}+p_{\rm 4~clumps}+p_{\rm >4~clumps})>0.5$.
